Simple connectome inference from partial correlation statistics in calcium imaging

### 2014-06-17 article.tex
discussion about the metric

At first sight, it appears on \textit{Table 1} that partial correlation
outperforms other methods. Two main reasons may explain this. First, the
quality of the infered network seems related to the conditioning part of the
method. In every method except with partial correlation, indirect links are
confounded with direct links.  Secondly, the gaussian assumption seems
verified and thus linear method appears to be competitive, e.g. the method
correlation already gives good results, with a non-linear method such as
GENIE3.

@@

@@

Directivity \textcolor{red}{appendix?} & 0.535 & 0.526 & 0.533 & 0.535 & 0.012 & 0.012 & 0.012 & 0.012 \\
Correlation            & 0.886 & 0.884 & 0.891 & 0.877 & 0.153 & 0.145 & 0.170 & 0.132 \\
GTE                    & 0.890 & 0.893 & 0.894 & 0.873 & 0.171 & 0.174 & 0.197 & 0.142 \\
GENIE3 (depth = 1)     & 0.878 & & & & 0.184 & & & \\
GENIE3 (depth = 2)     & 0.889 & & & & 0.217 & & & \\
GENIE3 (depth = 3)     & & & & & & & & \\
Partial correlation    & 0.930 &  0.927 &  0.926 &  0.923& 0.363  & 0.347 &  0.346 & 0.328 \\
Partial correlation (opt.) & 0.932 & 0.931 & 0.930 & 0.928 & 0.364 & 0.366 & 0.359 & 0.344 \\
Our approach (opt. \& avg.)    & 0.943 & 0.942 & 0.942 & 0.939 & 0.403 & 0.404 & 0.398 & 0.388 \\
Our approach + dir  \textcolor{red}{appendix?}   & 0.944 & 0.943 & 0.942 & 0.940 & 0.404 & 0.405 & 0.399 & 0.389 \\
\bottomrule
\end{tabular}
\end{table}

>>>>

At first sight, it appears on Tab.\ref{tab:tab1} that partial
correlation outperforms other methods. Two main reasons may explain this.
First, the quality of the infered network seems related to the conditioning
part of the method. In every method except with partial correlation, indirect
links are confounded with direct links.  Secondly, the gaussian assumption
seems verified and thus linear method appears to be competitive, e.g. the
method correlation already gives good results, with a non-linear method such
as GENIE3.\\

@@

However, the metrics used in the contest, i.e. the ROC AUC, does not seem to be consistent with the goal of the challenge, and thus to assess the quality of the \textbf{directed} infered network. Observing Tab.\ref{tab:tab1}, it appears that best methods only infer the \textbf{undirected} network by giving the same weights to both edges $(i,j)$ and $(j,i)$.

Indeed, if we consider that the infered network is the true network undirected, i.e. if for all links $(i,j)$, back links $(j,i)$ are added, then we get a very high score (about $0.9956$) according to this metric. Therefore, if we want to maximize the ROC AUC score, we only need to detect the links, no matter their direction.  Doing that, we only give up a gain margin of about $0.005\%$ which is not really interesting in the context of this challenge. 
We think that AUC PRC is more adapted because if we perform the same experiment, we get a score of about $0.7897$. In this case, the gain margin is about $0.2103$ which is much more interesting.

@@

Directivity \textcolor{red}{appendix?} & 0.535 & 0.526 & 0.533 & 0.535 & 0.012 & 0.012 & 0.012 & 0.012 \\
Correlation            & 0.886 & 0.884 & 0.891 & 0.877 & 0.153 & 0.145 & 0.170 & 0.132 \\
GENIE3 (depth = 1)     & 0.878 & & & & 0.184 & & & \\
GENIE3 (depth = 2)     & 0.889 & & & & 0.217 & & & \\
GENIE3 (depth = 3)     & & & & & & & & \\
GTE                    & 0.890 & 0.893 & 0.894 & 0.873 & 0.171 & 0.174 & 0.197 & 0.142 \\
Partial correlation    & 0.930 &  0.927 &  0.926 &  0.923& 0.363  & 0.347 &  0.346 & 0.328 \\
Partial correlation (opt.) & 0.932 & 0.931 & 0.930 & 0.928 & 0.364 & 0.366 & 0.359 & 0.344 \\
Our approach (opt. \& avg.)    & 0.943 & 0.942 & 0.942 & 0.939 & 0.403 & 0.404 & 0.398 & 0.388 \\
Our approach + dir  \textcolor{red}{appendix?}   & 0.944 & 0.943 & 0.942 & 0.940 & 0.404 & 0.405 & 0.399 & 0.389 \\
\bottomrule
\end{tabular}
\label{tab:tab1}
\end{table}
